---
title: "simpleRegress"
author: "Eileen Dunn"  
date: "Summer 2024"  
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## make a simple regression model

```{r}
bears <- read.csv(file="BEARS.csv",header = TRUE)
#adjust factors to be factors
bears$SEX<- as.factor(bears$SEX)
bears$MONTH<- as.factor(bears$MONTH)
```

## Model LENGTH ~ NECK  this is an example  

```{r}
simp <- lm(LENGTH~NECK,data=bears)
plot(LENGTH~NECK,data=bears)
abline(simp)
summary.lm(simp)
```

Adjusted r^2 = 84 this is better than the r^2 of LENGTH~NECK, this means that model LENGTH~HEADLEN is better.  

## Compare all models that predict LENGTH other than NECK. Then order them from worst to best by its value of Adjusted-Rsquared (bigger is better), or by RSE (residual standard error) where smaller is better. 

## length based on age  

```{r}
ageModel <- lm(LENGTH~AGE,data=bears)
plot(LENGTH~AGE,data=bears)
abline(ageModel)
summary.lm(ageModel)
```

adjusted $r^2$ for the age model is 0.51, thi sis less than the adjusted $r^2$ for neck, so this model is not as good as the neck model.  

## length based on chest  


```{r}
chestModel <- lm(LENGTH~CHEST,data=bears)
plot(LENGTH~CHEST,data=bears)
abline(chestModel)
summary.lm(chestModel)
```

adjusted r^2 = 0.79
The adjusted r^2 for headlength is 0.84, this shows that model LENGTH~HEADLEN is better than model LENGTH~CHEST  

## length based on sex  


```{r}
sexModel <- lm(LENGTH~SEX,data=bears)
plot(LENGTH~SEX,data=bears)
summary.lm(sexModel)
```

The adjusted r^2 = -0.0084 adjusted r^2 for LENGHTH~HEADLEN is better than LENGTH~SEX  

## length based on headlength  


```{r}
headlenModel <- lm(LENGTH~HEADLEN,data=bears)
plot(LENGTH~HEADLEN,data=bears)
abline(headlenModel)
summary.lm(headlenModel)
```

adjusted r^2= 53
the adjusted r^2 for headlen is better than headwith LENGTH~HEADLEN is better than LENGTH~HEADWTH


## length based on headwidth  


```{r}
heads <- lm(LENGTH~HEADWTH,data=bears)
plot(LENGTH~HEADWTH,data=bears)
abline(heads)
```
```{r}
summary.lm(heads)
```

adjusted r^2 = 53
the adjusted r^2 for HEADLEN is better than HEADWTH, LENGTH~HEADLEN is better than LENGTH~HEADWTH  

## lentgth based on weight  


```{r}
weight <- lm(LENGTH~WEIGHT,data=bears)
plot(LENGTH~WEIGHT,data=bears)
abline(weight)
```
```{r}
summary.lm(weight)
```

The adjusted r^2 = 0.74  
the adjusted r^2 for LENGTH~HEADLEN is 0.84  
this shows that LENGTH~HEADLEN is better the LENGTH~WEIGHT  

## Conclusion  

I conclusion the best model is LENGTH~HEADLEN is the best model it had the highest Adjusted r^2 which was 0.84 and all other models had lower scores.  